The speed of PyPy

### 2010-02-15 talk/pycon2010/pypyspeed/talk.tex
start working on talk

\setbeamerfont{big}{size*={64pt}{0pt}}

\title{The status of PyPy}

@@

\begin{frame}
  \frametitle{Introduction}
  \begin{itemize}
    \item What we did in the last year?
    \item What we achieved?
    \item Plans for near future
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Last year highlights}
  \pause
  \usebeamerfont{big}
  JIT
\end{frame}

\begin{frame}
  \frametitle{February 2009}
  \begin{tikzpicture}[remember picture, overlay]
    \node [shift={(1cm, -4cm)}] at (current page.north west) {
      \begin{tikzpicture}[remember picture, overlay, font={\fontsize{8pt}{0pt} \ttfamily}]
        \node [shift={(-1cm, 1.4cm)}, above right] {
          \includegraphics[width=.15\textwidth]{../common/pypy-logo.png}};
        \draw [thick, ->] (1.2cm, 2.5cm) -- (10cm, 2.5cm);
      \end{tikzpicture}
    };
  \end{tikzpicture}
  \begin{itemize}
    \item we merge experiments into mainline PyPy
      \pause
    \item compiles assembler, slows things down
  \end{itemize}
\end{frame}

\begin{frame}[fragile]
  \frametitle{March 2009}
  \begin{tikzpicture}[remember picture, overlay]
    \node [shift={(1cm, -4cm)}] at (current page.north west) {
      \begin{tikzpicture}[remember picture, overlay, font={\fontsize{8pt}{0pt} \ttfamily}]
        \node [shift={(-1cm, 1.4cm)}, above right] {
          \includegraphics[width=.15\textwidth]{../common/pypy-logo.png}};
        \draw [thick, ->] (1.2cm, 2.5cm) -- (10cm, 2.5cm);
        \draw [fill=blue!20,very thin] (1.7cm, 2.5cm) circle (.06cm);
        \draw (1.7cm, 2.44cm) -- (1.7cm, 2.0cm);
        \draw (1.5cm, 1.5cm) node [rotate=45] {JIT merge};  
      \end{tikzpicture}
    };
  \end{tikzpicture}
  \begin{itemize}
    \item first python example works!
      \begin{verbatim}
i = 0
while i < 1000000:
  i = i + 1
      \end{verbatim}
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{September 2009}
  \begin{tikzpicture}[remember picture, overlay]
    \node [shift={(1cm, -4cm)}] at (current page.north west) {
      \begin{tikzpicture}[remember picture, overlay, font={\fontsize{8pt}{0pt} \ttfamily}]
        \node [shift={(-1cm, 1.4cm)}, above right] {
          \includegraphics[width=.15\textwidth]{../common/pypy-logo.png}};
        \draw [thick, ->] (1.2cm, 2.5cm) -- (10cm, 2.5cm);
        \draw [fill=blue!20,very thin] (1.7cm, 2.5cm) circle (.06cm);
        \draw (1.7cm, 2.44cm) -- (1.7cm, 2.0cm);
        \draw (1.5cm, 1.5cm) node [rotate=45] {JIT merge};
        \draw [fill=blue!20,very thin] (4.3cm, 2.5cm) circle (.06cm);
        \draw (4.3cm, 2.44cm) -- (4.3cm, 2.0cm);
        \draw (3.95cm, 1.5cm) node [rotate=45] {first example};  
      \end{tikzpicture}
    };
  \end{tikzpicture}
  \begin{itemize}
    \item we manage to speed first non-trivial benchmark
    \item {\bf richards} - a simple, object oriented code
    \item a variety of optimizations involved
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{February 2010 - now}
  \begin{tikzpicture}[remember picture, overlay]
    \node [shift={(1cm, -4cm)}] at (current page.north west) {
      \begin{tikzpicture}[remember picture, overlay, font={\fontsize{8pt}{0pt} \ttfamily}]
        \node [shift={(-1cm, 1.4cm)}, above right] {
          \includegraphics[width=.15\textwidth]{../common/pypy-logo.png}};
        \draw [thick, ->] (1.2cm, 2.5cm) -- (10cm, 2.5cm);
        \draw [fill=blue!20,very thin] (1.7cm, 2.5cm) circle (.06cm);
        \draw (1.7cm, 2.44cm) -- (1.7cm, 2.0cm);
        \draw (1.5cm, 1.5cm) node [rotate=45] {JIT merge};
        \draw [fill=blue!20,very thin] (4.3cm, 2.5cm) circle (.06cm);
        \draw (4.3cm, 2.44cm) -- (4.3cm, 2.0cm);
        \draw (3.95cm, 1.5cm) node [rotate=45] {first example};  
        \draw [fill=blue!20,very thin] (7.3cm, 2.5cm) circle (.06cm);
        \draw (7.3cm, 2.44cm) -- (7.3cm, 2.0cm);
        \draw (6.95cm, 1.5cm) node [rotate=45] {richards};  
      \end{tikzpicture}
    };
  \end{tikzpicture}
  We run a variety of benchmarks nightly. Pictures!
\end{frame}

\begin{frame}
  \frametitle{Status}
  \begin{block}{Status of today}
    \begin{itemize}
    \item JIT is fairly good
      \pause
    \item can consume a lot of memory
      \pause
    \item can recompile code too many times
      \pause
    \item GCs are pretty good
      \pause
    \item runtime algorithms are not as polished as CPython
      ones (eg strings, regexes)
    \end{itemize}
  \end{block}
\end{frame}

\begin{frame}
  \frametitle{Near future}
  \begin{itemize}
    \item release March 2010
      \pause
    \item first PyPy release to contain a working JIT!
    \item mostly to try it out
      \pause
    \item we just got a bit more {\bf EU funding}
    \item we're actively looking for {\bf sponsoring} to continue PyPy development
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Come to my next talk!}
  \begin{block}{speed of PyPy}
    ... or how to write programs that play well with the JIT
  \end{block}
\end{frame}

\begin{frame}
  \frametitle{That's all}
  Thank you for your attention
  \begin{itemize}
    \item http://morepypy.blogspot.com
    \item http://pypy.org
    \item http://merlinux.eu
    \item these slides are already online:
    \item http://tinyurl.com/pypyslides
  \end{itemize}
\end{frame}

>>>>

\title{The speed of PyPy}

@@

\begin{frame}
  \frametitle{JIT - what's that about?}
  \begin{itemize}
    \item {\bf JIT is not a magical device!}
      \pause
    \item removes bytecode overhead
    \item removes frame overhead
    \item can make runtime decisions
    \item more classic optimization that can follow
\end{frame}

\begin{frame}
  \frametitle{The main idea}
  \begin{itemize}
    \item python has advanced features (frame introspection,
      arbitrary code execution, overloading globals)
    \item with JIT, you don't pay for them if you don't use
      them
    \item however, you pay if you use them, but they work
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{A piece of advice}
  \begin{itemize}
    \item Don't use advanced features if you don't have to
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{xxx}
\end{frame}
